Increased productivity using Atomify - a real-time LAMMPS visualizer

### 2018-06-01 ms.tex
Wrote more intro

\documentclass[aps,pre,twocolumn,letterpaper,floatfix,nofootinbib]{revtex4}
\usepackage{graphicx} 
\usepackage{amsmath,amssymb,amsfonts} 
\usepackage{mathtools}
\usepackage{pdfpages}
\usepackage{afterpage}
\usepackage[hidelinks]{hyperref} 
\usepackage{epstopdf}
\usepackage{todo}
\usepackage{menukeys}

\begin{document}

@@

\section{Introduction}
%
Molecular dynamics (MD) is a common technique for simulating the motion of atoms,
molecules and larger particles in a broad range of fields such as physics, chemistry and biology \citep{van2001reaxff, allen2017computer, }.
% TODO does MD also include larger particles?
% TODO which fields?
Over the years, increasingly sophisticated methods have been developed.
% TODO what methods?
With the large number of force fields, advanced techniques and GPU acceleration,
implementing a good molecular dynamics simulation is a challenging task.
There are many software packages for molecular modelling
available\footnote{\url{https://en.wikipedia.org/wiki/Comparison_of_software_for_molecular_mechanics_modeling} compares different softwares for molecular modeling.}.
Some packages, such as LAMMPS\citep{Plimpton1995Fast}, Gromacs\citep{berendsen1995gromacs}
and NAMD\citep{Phillips2005Scalable} have been developed for more than two decades 
and contain more than one million lines of code each.

The work in this paper is focused on LAMMPS.
Although the software is well-documented and enables the user to perform
advanced simulations, it can be demanding to learn how to use it well.
A LAMMPS simulation is performed by a series of commands which are executed one by
one.
Physical quantities such as energy, temperature and stress can be measured by
adding \textit{computes}.
The time evolution of these quantities can in turn be saved to file for plotting
and further analysis.
Particle trajectories can be saved to a file and visualized by tools such as
VMD\citep{Humphrey1996Vmd} or OVITO\citep{Stukowski2009Visualization}.
The lack of an intuitive GUI and visualization\footnote{GUI is listed as a non-feature on the web page:
\url{http://lammps.sandia.gov/non_features.html}.} makes the workflow somewhat
complicated, see figure \ref{fig:flowchart}.

@@

\section{Features}

The main features of Atomify are script editing,
live visualization, and live plotting of computed quantities.
In addition, Atomify comes packaged with multiple examples,
many of which have been made by members of the LAMMPS community.
These examples are easily explored within Atomify.

>>>>

\documentclass[aps,pre,twocolumn,letterpaper,floatfix,nofootinbib]{revtex4}
\usepackage{graphicx} 
\usepackage{amsmath,amssymb,amsfonts} 
\usepackage{mathtools}
\usepackage{pdfpages}
\usepackage{afterpage}
\usepackage[hidelinks]{hyperref} 
\usepackage{epstopdf}
\usepackage{todo}
\usepackage{menukeys}
\usepackage{siunitx}
\begin{document}

@@

\section{Introduction}
Our nature is a very complex system with complicated rules that determine how and where mass, energy and momentum is transferred.
Many macroscopic observable effects result from microscopic details on the atomic level such as melting of ice, and even life.
Understanding the world from a bottom-up perspective requires us to be able to model a physical system from the atomic scale where the timescale is of order femtoseconds.
Many physical and chemical processes happen on nanosecond timescales, some even during a few picoseconds, which means thousands or millions of timesteps of \SI{1}{\femto\second} need to be calculated.
Much has happened since the first molecular dynamics simulation of hard spheres in the 1950s\citep{alder1957phase,alder1959studies}, and later, in 1964, a more realistic system with the Lennard Jones potential studying diffusion in argon\citep{rahman1964correlations}.
The computing power has seen a tremendous increase from the CDC 3600 used in \citep{rahman1964correlations}, which performed less than 1 million instructions per second whereas modern 3 Ghz processors in typical laptops can do about 100 billion instructions per seconds.
With this rapidly increase of computing power, the lengthscales and timescales achievable in numeriments\footnote{Numeriment is a numerical experiment, or a simulation on the computer.} is finally catching up with some experiments (ref).

In molecular dynamics, a set of atoms is integrated using Newton's second law in a classical force field so the full phase space is known at all times and statistical properties can be calculated from averages assuming ergodicity.
It has been prooved successful in reproducing and explaining multiple phenomenon such as mechanical properties \citep{jiang2009young, campbell1999structural, ning2012mechanical}, chemical reactions \citep{van2001reaxff, brenner2002second}, liquid properties\citep{allen2017computer}.
Also in biology, molecular dynamics is useful blabla.

Over the years, increasingly sophisticated methods have been developed.
For instance, a non-equilibrium method to measure thermal conductivity using heat exchange, was recently improved\citep{wirnsberger2015enhanced} to conserve energy better.
To study first order phase transitions, methods like the generalized replica exchange method (gREM)\citep{kim2010generalized} can be used and nudge elastic band (NEB)\citep{henkelman2000climbing, henkelman2000improved} is often used to study energy barriers in for instance chemical reactions.
In addition, a huge amount of force fields exists, each being good at different things.
New hardware and a sea of optimization tricks makes implementing a good molecular dynamics code a challenging task, not to mention maintaining it over time on new arcitechtures.
Each new feature or technique like the ones metioned above, may take days, if not weeks, to develop, preventing researchers to work on their scientific projects.

Thanksfully, there already exists several software packages for atomistic simulations, many of which have been developed for more than two decades\footnote{\url{https://en.wikipedia.org/wiki/Comparison_of_software_for_molecular_mechanics_modeling} compares different softwares for molecular modeling.}.
Some of these packages, such as LAMMPS\citep{Plimpton1995Fast}, Gromacs\citep{berendsen1995gromacs}, NAMD\citep{Phillips2005Scalable} and CHARMM\citep{brooks2009charmm}, are open source and being actively developed with hundreds of contributers.
%After the rise of powerful graphical processing units (GPU's), most of these codes eventually added support of using accelerators\citep{brown2011implementing}.

The work in this paper is focused on LAMMPS.
Although the software is well-documented\footnote{The documentation of LAMMPS is known for being very detailed precise.} and enables the user to perform
advanced simulations, it can be demanding to learn how to use it well.
LAMMPS has a huge list of implemented features which makes it very flexible with its own scripting language.
A LAMMPS simulation is performed by a series of commands which are executed one by one.

Physical quantities such as energy, temperature and stress can be measured by adding \textit{computes}.
The time evolution of these quantities can in turn be saved to file for plotting
and further analysis.
Particle trajectories can be saved to a file and visualized by tools such as
VMD\citep{Humphrey1996Vmd} or OVITO\citep{Stukowski2009Visualization}.
The lack of an intuitive GUI and visualization\footnote{GUI is listed as a non-feature on the web page:
\url{http://lammps.sandia.gov/non_features.html}.} makes the workflow somewhat
complicated, see figure \ref{fig:flowchart}.

@@

\section{Features}
The main features of Atomify are script editing,
live visualization, and live plotting of computed quantities.
In addition, Atomify comes packaged with multiple examples,
many of which have been made by members of the LAMMPS community.
These examples are easily explored within Atomify.
